---
title: "Kaggle Titanic"
output: html_document
---

This is the Kaggle Titanic competition, which is a learning example
rather than a real competition.

### Data Munging

The major portion of the task here will be cleaning and imputing
the input data.  So, start by reading in the training data:
```{r}
### Reads titanic data
read.Titanic <- function(path.name, file.name) {
    filename <- paste(path.name, file.name, sep="/")
    col.types <- c('integer',   # PassengerId
                    'factor',    # Survived 
                    'factor',    # Pclass
                    'character', # Name
                    'factor',    # Sex
                    'numeric',   # Age
                    'integer',   # SibSp
                    'integer',   # Parch
                    'character', # Ticket
                    'numeric',   # Fare
                    'character', # Cabin
                    'factor')     # Embarked
    read.csv(filename, colClasses=col.types,
             na.strings=c("NA", ""))
}
train.data <- read.Titanic('.', 'train.csv')
```

Now, there are some missing data that we need to impute.  A huge fraction
of the cabin numbers are missing, which means it may be necessary to throw
them out.  A good 20% of the ages are missing -- perhaps it is possible to
impute those.  And a tiny number of embarked values are missing.  Since the
majority of the passengers embarked at Southampton -- even those in first
class, which the two missing values are from -- it's should be okay to 
just assign those to the 'S' class.

What about age?  It probably correlates with class:
```{r}
boxplot(train.data$Age ~ train.data$Pclass, main="Age vs. Passenger Class",
        xlab="Passenger Class", ylab="Age [yrs]")
```
So that's probably useful.

Another useful bit of information may come from the passenger names.